---
title: "SLC35A2 Olig2 cKO snRNAseq analysis"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
# Project set up

## Random Seed
```{r Setting seed, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}}
set.seed(0827)
```
## Library load
```{r Library loading, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
#### Library load ####
# Analysis
library(future)
library(Seurat)
library(glmGamPoi)
library(SeuratExtend)
library(clusterProfiler)
library(DoubletFinder)
library(parallel)
library(future)

# library(rrvgo)
library(SoupX)
library(miloR)
library(SingleCellExperiment)
library(scater)
library(scran)

# Stat test
library(presto)
library(speckle)
library(scProportionTest)

# Reference
library(org.Mm.eg.db)
library(BSgenome.Mmusculus.UCSC.mm10)

# File load and save
library(qs)

# Load HDF5 library from IGM
# dyn.load("/igm/apps/hdf5/hdf5-1.12.1/lib/libhdf5_hl.so.200")  
library(hdf5r)

# Data manipulation
library(tidyverse)

# Visualization
library(ggplot2)
library(ggstats)
library(patchwork)
library(EnhancedVolcano)
library(enrichplot)
library(ggtext)
```

```{r}
options(Ncpus = parallel::detectCores()-1)
options(future.globals.maxSize = 16000 * 1024^2)
```

## Source Code list
```{r Source Code list, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Load custom function scripts
source("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/01__load_rna_from_h5.R")
source("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/02__create_seurat_object.R")
source("~/SLC35A2_Olig2cKO_snRNA_JY_JY/WIP/SCRIPTS/SourceCode/03__rna_qc.R")
source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/04__process_rna.R')
source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/05__rpca_rna_integration.R')
source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/06__process_integrated.R')
source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/07__clustering.R')
source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/08__ref_mousebrain_annotation.R')
```

## Folder directory
```{r directory list, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Directory setup
h5_input_dir <- 
  '/igm/projects/250718_GSL-EM-4483'
save_dir <- 
  '~/SLC35A2_Olig2cKO_snRNA_JY/WIP'
qsave_dir <- 
  file.path(save_dir, "QSAVE")
qc_dir    <- 
  file.path(save_dir, "QC")
csv_dir <- 
  file.path(save_dir, 'CSV')
plot_dir <- 
  file.path(save_dir, 'PLOTS')

# Create folders if needed
dir.create(save_dir, recursive = TRUE)
dir.create(qsave_dir, recursive = TRUE)
dir.create(qc_dir, recursive = TRUE)
dir.create(csv_dir, recursive = TRUE)
dir.create(plot_dir, recursive = TRUE)
```

## Sample list set
```{r Sample list setup, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Set sample list name = index, item = samples
samples_named <- 
  list(
    "FF061" = c("SOM333", "SOM318", "SOM303", "SOM301"),
    "FF062" = c("SOM326", "SOM325", "SOM323", "SOM322"),
    "FF063" = c("SOM329", "SOM328", "SOM327", "SOM302"),
    "FF064" = c("SOM378", "SOM377", "SOM376", "P07"))

# Set sample list without names
samples <- 
  names(seurat_list)

# Sample Info
info <- list(
  "SOM301" = c("P23", "cKO"),
  "SOM303" = c("P23", "cKO"),
  "SOM318" = c("P08", "cKO"),
  "SOM333" = c("P23", "Control"),
  "SOM322" = c("P56", "Control"),
  "SOM323" = c("P56", "Control"),
  "SOM325" = c("P56", "cKO"),
  "SOM326" = c("P56", "cKO"),
  "SOM302" = c("P23", "Control"),
  "SOM327" = c("P08", "Control"),
  "SOM328" = c("P08", "Control"),
  "SOM329" = c("P08", "cKO"),
  "SOM376" = c("P00", "cKO"),
  "SOM377" = c("P00", "Control"),
  "SOM378" = c("P00", "Control"),
  "P07" = c("P00", "cKO")
)
```

# Initial processing

## Load data files
```{r [O] Load RNA data from H5 files, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/01__load_rna_from_h5.R")
rna_list <- 
  load_rna_from_h5(h5_input_dir,
                   samples_named,
                   save = T,
                   save_dir = qsave_dir)

# Save
qs::qsave(rna_list, 
          file = "~/SLC35A2_Olig2cKO_snRNA_JY/WIP/QSAVE/01_rna_list.qs")

```

## Seurat Object creation
```{r [O] Create Seurat Object from RNA data, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/02__create_seurat_object.R")

# load previous object if needed
# rna_list <- qs::qread("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/QSAVE/01_rna_list.qs")

seurat_list <- 
  create_seurat_object.R(rna_list,
                       h5_input_dir, # for meta data file path
                       samples_named,
                       project = 'SLC35A2Olig',
                       min.cell = 3,
                       min.feature = 200,
                       save = T,
                       save_dir = qsave_dir)

# Raw count matrices loaded
# Seurat objects created with  metadata
# Sample/index identifiers added
# Mitochondrial gene percentages calculated

# Save
qs::qsave(seurat_list, 
          file = file.path(qsave_dir, "02_rna_list_seurat.qs"))
```

## Initial Vizualization
```{r [V] Initial Visual Overview , echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# R script located at "~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/02_Initial_VlnPlot.R"
plots <- qc_VlnPlot(seurat_list, save_dir = plot_dir)
```

## Basic metrics
```{r [S] Initial process - Stats Table, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# From the literature
# Cell-Level Filtering:
## UMI counts: 1000 < UMI < 25,000
## Gene features: > 400 genes (some use > 200)
## Mitochondrial %: < 15%
# Gene-Level Filtering:
## Min cells expressing: ≥ 3 cells
## Remove mitochondrial genes for downstream analysis

# Create table with counts information
SummaryTable <- data.frame()

for (sample in seurat_list) {
  dt <- data.frame(
    # Sample name
    MouseID = unique(sample$MouseID),
    
    TotalCells = nrow(sample@meta.data),
    
    nCount_mean = mean(sample$nCount_RNA),
    nFeature_RNA_mean = mean(sample$nFeature_RNA),
    percent_mt_rna_mean = mean(sample$percent_mt_rna),
    
    nCount_median = median(sample$nCount_RNA),
    nFeature_RNA_median = median(sample$nFeature_RNA),
    percent_mt_rna_median = median(sample$percent_mt_rna),
    
    nCount_min = min(sample$nCount_RNA),
    nCount_max = max(sample$nCount_RNA),
    
    nFeature_RNA_min = min(sample$nFeature_RNA),
    nFeature_RNA_max = max(sample$nFeature_RNA),
    
    percent_mt_rna_min = min(sample$percent_mt_rna),
    percent_mt_rna_max = max(sample$percent_mt_rna)
  )
  
  # Put all in one table
  SummaryTable <- bind_rows(SummaryTable, dt)
}

write.csv(SummaryTable, file = file.path(csv_dir, "Initial_stats.csv"))

#### Same table but using map_dfr  #### 
SummaryTable <- map_dfr(names(seurat_list), ~{
  meta <- seurat_list[[.x]]@meta.data
  data.frame(
    # Sample name
    MouseID = unique(meta$MouseID),

    TotalCells = nrow(meta),

    nCount_mean = mean(meta$nCount_RNA),
    nFeature_RNA_mean = mean(meta$nFeature_RNA),
    percent_mt_rna_mean = mean(meta$percent_mt_rna),

    nCount_median = median(meta$nCount_RNA),
    nFeature_RNA_median = median(meta$nFeature_RNA),
    percent_mt_rna_median = median(meta$percent_mt_rna),

    nCount_min = min(meta$nCount_RNA),
    nCount_max = max(meta$nCount_RNA),

    nFeature_RNA_min = min(meta$nFeature_RNA),
    nFeature_RNA_max = max(meta$nFeature_RNA),

    percent_mt_rna_min = min(meta$percent_mt_rna),
    percent_mt_rna_max = max(meta$percent_mt_rna))
})

```

## QC metric exprlore
```{r [S] Initial process - QC threshold test, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/03_qc_threshold_test.R')
QC_test <- qc_threshold(seurat_list = seurat_list,
                         min_count = 1000,
                         max_count = 25000,
                         min_feature = 400,
                         max_mt = 5)
```

# Filtering and Processing
## Filter based on QC matrics
```{r [O] Initial process - Filter, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/03__rna_qc.R")

# load previous object if needed
# seurat_list <- qs::qread("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/QSAVE/02_rna_list_seurat.qs")

# Filter the cells
filtered_list <- rna_qc(
  seurat_list,
  samples,
  nCount_RNA_low = 1000,
  nCount_RNA_high = 25000,
  nFeature_RNA_low = 400,
  mt_rna = 5,
  save = T,
  save_dir_qs = qsave_dir,
  save_dir_qc = qc_dir)

# Check filtered cell counts
QC_summary <- read_csv("WIP/QC/QC_summary.csv")

# Save
qs::qsave(filtered_list, 
          file = file.path(qsave_dir, "03_filtered_rna_list.qs"))

```

## Pre-processing
```{r [O] Normalization and preprocessing, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/04__process_rna.R")

# Read previous object if needed.
# filtered_list <- qs::qread(file.path(qsave_dir, "03_filtered_rna_list.qs"))

# Normalization, SCTrancform, PCA and UMAP/TSNE
processed_list <- process_rna(
  filtered_list,
  save = TRUE,
  save_dir = qsave_dir)

# Save
qs::qsave(processed_list, 
          file = file.path(qsave_dir, '04_processed_rna_list.qs'))
```

# Integration
Integration with SCT method -> failed due to limit of data size can handeled by Seurat

## Integration with RPCA method
Integration with SCT can't cover the number of cells in our object so trying RPCA method in next section
```{r [O] Integrate RNA objects with RPCA, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/05__rpca_rna_integration.R')

# Read previous object if needed.
# processed_list <- qs::qread(file.path(qsave_dir, "04_processed_rna_list.qs"))

# Integration
integrated <- rpca_integration(processed_list, 
                               save = TRUE,
                               save_dir = qsave_dir)

```

# Process Integrated object
## PCA, Dimentional Reduction
```{r [O] Process integrated object, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/06__process_integrated.R')

# Read previous object if needed.
# integrated <- qs::qread(file.path(qsave_dir, "05_integrated_rna.qs"))

# Process integrated object
integrated <- process_integrated(integrated,
                                 save = TRUE,
                                 qsave_dir = qsave_dir)

```

### Initial visualization - DimPlot
```{r [V] Initial DimPlot after integration, message=TRUE, warning=TRUE, paged.print=TRUE}
# set assay and Idents
Seurat::DefaultAssay(integrated) <- "RNA"
Seurat::Idents(integrated) <- "MouseID"

# DimPLot
dp <- Seurat::DimPlot(integrated, reduction = 'umap', group.by = 'Timepoint', alpha = 0.3) |
  Seurat::DimPlot(integrated, reduction = 'umap', group.by = 'Genotype', alpha = 0.3) |
  Seurat::DimPlot(integrated, reduction = 'umap', group.by = "MouseID", alpha = 0.3) |
  Seurat::DimPlot(integrated, reduction = 'umap', group.by = "SampleID", alpha = 0.3)

# Save
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_initial_all.png"), dp, width = 18, height = 4, dpi = 300)

Seurat::DimPlot(integrated, reduction = 'umap', split.by = 'Timepoint', alpha = 0.3)
Seurat::DimPlot(integrated, reduction = 'umap', split.by = 'Genotype', alpha = 0.3)
Seurat::DimPlot(integrated, reduction = 'umap', split.by = "MouseID", alpha = 0.3)
Seurat::DimPlot(integrated, reduction = 'umap', split.by = "SampleID", alpha = 0.3)
```

## Cluster and Markers, additional meta data
```{r [O] Clusteriazation, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/07__clustering.R')

# Read previous object if needed.
# integrated <- qs::qread(file.path(qsave_dir, "06_integrated_rna.qs"))

integrated <- clustering(integrated,
                         range_start = 0.4,
                         range_end = 1.2,
                         range_step = 0.1,
                         save = TRUE,
                         qsave_dir = qsave_dir,
                         markers_dir = file.path(csv_dir, "Markers"))


# Additional meta data
# Forgot to add this early in the process so adding now.
# Convert info list to lookup vectors
timepoint_lookup <- sapply(info, function(x) x[1])
genotype_lookup <- sapply(info, function(x) x[2])

# Add to metadata
integrated@meta.data <- integrated@meta.data %>%
  mutate(
    TimePoint = timepoint_lookup[MouseID],
    GenoType = genotype_lookup[MouseID]
  )

# Add additional meta data combine age and genotype
integrated@meta.data$TimeGeno <- paste0(integrated@meta.data$TimePoint, "_", integrated@meta.data$GenoType)

# Join layers after integration in RNA assay
Seurat::DefaultAssay(integrated) <- "RNA"
integrated <- SeuratObject::JoinLayers(integrated)

# Save
qs::qsave(integrated, file = file.path(qsave_dir, "08_clustered_metadata_added.qs"))
```

## Use Mousebrain.org object as reference for cell typing. Using Jesse's custom object.
```{r [O] Cell type, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/08_ref_mousebrain_annotation.R')

# Read previous object if needed.
# integrated <- qs::qread(file.path(qsave_dir, "08_clustered_metadata_added.qs"))

# Read reference object for cell typing
ref_obj <- readRDS("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/mouseatlas_cut_processed.rds")

Seurat::DefaultAssay(integrated) <- "SCT"
Seurat::DefaultAssay(ref_obj) <- "SCT"

# run mouse annotation
integrated <- ref_mousebrain_annotation(ref_obj,
                                        integrated,
                                        save = T,
                                        save_dir = qsave_dir)

# Also, cleaned out metadat. all prediction score were deleted.
```

```{r [O] Clean ref annotation, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object if needed.
# integrated <- qs::qread(file.path(qsave_dir, "09_ref_annotated_obj.qs"))

# Ref cell type prediction will be stored in oligsub_man2$predicted.id.
Seurat::Idents(integrated) <- integrated$predicted.id

# Visualize
Seurat::DimPlot(integrated, reduction = "umap", label = T, repel = T) + NoLegend()

# get names and shorten them
ln <- unique(integrated$predicted.id)

sn <- c("Ex_Cortex", "Microglia", "Ex_CA3", "NPC", "Ex_CA1", "Interneuron", "Astrocytes", "Granule_DG", "Microglia",
        "MFOL", "Vascular", "Interneuron", "COP", "Interneuron", "Interneuron", "CGE", "Ependymal", "CGE",
        "Mature_OL", "OPC", "Cajal_Retzius", "Interneuron", "Pericytes", "Interneuron", "RGL_DG", "Astrocytes", "Granule_Blast_DG",
        "Macrophages", "Vascular","MGE", "NFOL", "Trilaminar", "Vascular", "Vascular", "Interneuron", "Vascular", "Interneuron")

# Mapping sn-ln
name_map <- setNames(sn, ln)

# Function to replace long names to short names
shortened <- function(name) {ifelse(name %in% names(name_map), name_map[name], name)}

# Apply function
integrated$ref_short <- shortened(integrated$predicted.id)

# Check UMAP
Seurat::Idents(integrated) <- "ref_short"
dp_ref <- Seurat::DimPlot(integrated, reduction = "umap", label = T, repel = T) + Seurat::NoLegend()

# Save
ggplot2::ggsave(dp_ref, file = file.path(plot_dir, "DimPlot_mousebrain_ref.png"), width = 8, height = 8, dpi = 300)
qs::qsave(integrated, file = file.path(qsave_dir, "09_ref_annotated_obj.qs"))
```

## Milo try
```{r [O] Milo for identifying cluster between cKO vs Control, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
## ~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/99_Milo.R
# https://www.nature.com/articles/s41587-021-01033-z
```

## Gene list
```{r [S] Gene list, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# "~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/99_gene_list.R"
```

## Manual Cell type check
```{r [V] Cell type check with marker genes, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# "~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/99_CellTypeWork.R"
Seurat::Idents(integrated) <- "CellTypes" 
Seurat::DimPlot(integrated, reduction = "umap", label = TRUE, repel = TRUE) + Seurat::NoLegend()
```

## SLC35A2 exrpression check
```{r [V] Check SLC35A2 expression, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Assay and idents set
Seurat::DefaultAssay(integrated) <- "SCT"
Seurat::Idents(integrated) <- "ref_short"

# Visualize
SeuratExtend::VlnPlot2(oligsub, features = "Slc35a2",show.mean = F, pt = F, box = F)
Seurat::FeaturePlot(integrated, features = "Slc35a2", label = F, repel = T, split.by = "TimeGeno")

# Save for refrence
ggplot2::ggsave(slc35a2, file = file.path(plot_dir, "Feature_slc35a2.png"), width = 32, height = 8, dpi = 300)

# Looking only Olig cells
oligsub <- subset(integrated, integrated$ref_short %in% c("OPC", "COP", "NFOL", "MFOL", "Mature_OL"))
oligsub$ref_short <- factor(oligsub$ref_short, levels = c("OPC", "COP", "NFOL", "MFOL", "Mature_OL"))

# Assay and idents set
Seurat::DefaultAssay(oligsub) <- "SCT"
Seurat::Idents(oligsub) <- "MouseID"
SeuratExtend::VlnPlot2(oligsub, features = "Slc35a2",show.mean = F, pt = F, box = F)
# Slc35a2 expression doesn't seem to be reduced in cKO cells. it might be due to ambient RNA from dissociateion/library prep. Plan to implement SoupX (Tracy will provide the script to start). Then I can compare once again. 
# SoupX didn't change much.
## Nuclear RNA only, 
```

## Cell count check
```{r [S] Cell count Summary, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
## "~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/99_CellCount_Slc35a2Check.R"
# Done cell count summary, Slc35a2 expressing cell check
# plot save in "./WIP/Plots/Cell_counts
```

## Visualize Cell tyep markers
```{r [V]] DotPlot, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object
# integrated <- qs::qread(file.path(qsave_dir, "10_metadata_edited_obj.qs"))

# Order Cell types
integrated$CellTypes <- factor(integrated$CellTypes, levels = c('Astrocytes', 'Microglia', 'Ependymal', 'ExNeurons', 'InNeurons', 'NPC', 'OPC', 'NFOL', 'MFOL', 'Pericytes'))

# Set marker genes to plot
markers <- c('Aqp4', 'Aldh1l1', 'Gfap', 'P2ry12', 'Aif1', 'Tmem212', 'Foxj1', 'Satb2',  'Gad1','Cux1', 'Sst', 'Ndnf', 'Nes', 'Pdgfra', 'Cspg4', 'Enpp6', 'Plp1', 'Mog', 'Vtn', 'Pdgfrb')

# Create DotPlot
dp <- Seurat::DotPlot(integrated, 
        features = markers, 
        group.by = "CellTypes",
        cols = c("darkgrey", "indianred1")) +
  ggplot2::coord_flip() +
  ggplot2::theme(
        axis.text.x =
            ggplot2::element_text(
                angle = 90,
                vjust = 0.8,
                hjust = 1,
                size = 10
            ),
        axis.text.y =
            ggplot2::element_text(
                size = 10
            ),
        axis.title.x =
            ggplot2::element_blank(),
        axis.title.y =
            ggplot2::element_blank(),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)
    )

# Additional DimPlot
dimp <- SeuratExtend::DimPlot2(integrated, 
                               reduction = 'umap', 
                               group.by = 'CellTypes',
                               label = TRUE,
                               label.size = 5,
                               repel = TRUE,
                               theme = list(
                                 Seurat::NoLegend(), 
                                 Seurat::NoAxes(), 
                                 ggtitle("UMAP by Cell Type")
                                 )) +  
  theme_umap_arrows(x_label = "UMAP_1",
                    y_label = "UMAP_2") 

# Save
ggplot2::ggsave(dp, filename = file.path(plot_dir, "DotPlot_CellMarkers.png"), width = 7, height = 7, dpi = 300)
ggplot2::ggsave(dimp, filename = file.path(plot_dir, "DimPlot_CellType.png"), width = 7, height = 7, dpi = 300)

```

# Doublet process
```{r [O] Doublet filter check, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Get individual RNA objects
processed_list <- qs::qread(file.path(qsave_dir, "04_processed_rna_list.qs"))
integrated <- qs::qread(file.path(qsave_dir, "10_metadata_edited_obj.qs"))

# Transfer cluster identity
meta <- integrated@meta.data %>%
  tibble::rownames_to_column("CellID") %>%
  select(CellID, ref_short, CellTypes)

for (i in seq_along(processed_list)) {
  obj_meta <- processed_list[[i]]@meta.data %>%
    tibble::rownames_to_column("CellID")
  
  new_meta <- obj_meta %>%
    left_join(meta, by = "CellID")%>%
    tibble::column_to_rownames("CellID")
  
  processed_list[[i]]@meta.data <- new_meta
    
}

# save
qs::qsave(processed_list, file = file.path(qsave_dir, "04_processed_rna_list_id.qs"))

# Run doubletfinder
source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/99_doublet_process_rna.R')
processed_list <- qs::qread(file.path(qsave_dir, "04_processed_rna_list_id.qs"))
integrated <- qs::qread(file.path(qsave_dir, "10_metadata_edited_obj.qs"))

# Run
processed_list <- find_dbl(processed_list, save = TRUE, save_dir = qsave_dir)

# Collect all metadata from processed_list
all_individual_meta <- list()

for (i in seq_along(processed_list)) {
  individual_meta <- processed_list[[i]]@meta.data %>%
    tibble::rownames_to_column("CellID") %>%
    select(CellID, contains("Doublet_Status")) # Select doublet-related columns
    
  all_individual_meta[[i]] <- individual_meta
}

# Combine all individual metadata
combined_individual_meta <- do.call(bind_rows, all_individual_meta)

# Transfer back to integrated object
integrated_meta <- integrated@meta.data %>%
  tibble::rownames_to_column("CellID") %>%
  left_join(combined_individual_meta, by = "CellID") %>%
  tibble::column_to_rownames("CellID")

# Update integrated object metadata
integrated@meta.data <- integrated_meta

# save
qs::qsave(integrated, file = file.path(qsave_dir, "11_dbl_added.qs"))
```

## Doublet counts
```{r [S] Doublet Stats, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# integrated <- qs::qread(file.path(qsave_dir, "11_dbl_added.qs"))
doublet_ratio <- round(table(integrated$Doublet_Status)[[1]]/table(integrated$Doublet_Status)[[2]],3)

doublet_grouped <- integrated@meta.data %>%
  group_by(ref_short) %>%
  count(Doublet_Status)
```

# (wrong approach- subsettied obj should be processed from integration) Work only with Olig population
## Subset Olig only and re-process
```{r [O] Subset, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object
# integrated <- qs::qread(file.path(qsave_dir, "10_metadata_edited_obj.qs"))

# Subset based on manual
oligsub_man <- subset(integrated, integrated$CellTypes %in% c("OPC", "NFOL", "MFOL"))

# Subset based on manual
oligsub_ref <- subset(integrated, integrated$ref_short %in% c("OPC", "COP", "NFOL", "MFOL", "Mature_OL"))

# Save
qs::qsave(oligsub_man, file = file.path(qsave_dir, "a1_OligOnlyInit_manual.qs"))
qs::qsave(oligsub_ref, file = file.path(qsave_dir, "a1_OligOnlyInit_refshort.qs"))
```   

```{r [O] Re-process, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object
# oligsub_man <- qs::qread(file.path(qsave_dir, "a1_OligOnlyInit_refshort.qs"))
# oligsub_ref <- qs::qread(file.path(qsave_dir, "a1_OligOnlyInit_manual.qs"))

# Normalize/process 
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/06_process_oligsub.R')

# Process integrated object
oligsub_man <- process_oligsub(oligsub_man,
                               save = TRUE,
                               suffix = "_manual",
                               qsave_dir = qsave_dir)

oligsub_ref <- process_oligsub(oligsub_ref,
                           save = TRUE,
                           suffix = "_refshort",
                           qsave_dir = qsave_dir)
```

## Initial visualization
### Manual cluster
```{r [V] DimPlot manual, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object
# oligsub_man <- qs::qread(file.path(qsave_dir, "a2_olig_processed_manual.qs"))

# set assay and Idents
Seurat::DefaultAssay(oligsub_man) <- "RNA"
Seurat::Idents(oligsub_man) <- "MouseID"

# DimPLot
dp <- Seurat::DimPlot(oligsub_man, reduction = 'umap', group.by = 'TimePoint', alpha = 0.3) |
  Seurat::DimPlot(oligsub_man, reduction = 'umap', group.by = 'GenoType', alpha = 0.3) |
  Seurat::DimPlot(oligsub_man, reduction = 'umap', group.by = "MouseID", alpha = 0.3) |
  Seurat::DimPlot(oligsub_man, reduction = 'umap', group.by = "SampleID", alpha = 0.3)

dp_time <- Seurat::DimPlot(oligsub_man, reduction = 'umap', group.by = 'TimePoint', alpha = 0.3, split.by = "GenoType")

dp_geno <- Seurat::DimPlot(oligsub_man, reduction = 'umap', group.by = 'GenoType', alpha = 0.3, split.by = "TimePoint")

dp_tg <- Seurat::DimPlot(oligsub_man, reduction = 'umap', split.by = 'TimeGeno', alpha = 0.3)

# Save
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_all_manual.png"), dp, width = 18, height = 4, dpi = 300)
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_timepoint_manual.png"), dp_time, width = 9, height = 4, dpi = 300)
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_genotype_manual.png"), dp_geno, width = 9, height = 4, dpi = 300)
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_TimeGeno_manual.png"), dp_tg, width = 18, height = 4, dpi = 300)
```

### ref_short cluster
```{r [V] DimPlot refshort, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object
# oligsub_ref <- qs::qread(file.path(qsave_dir, "a2_olig_processed_refshort.qs"))

# set assay and Idents
Seurat::DefaultAssay(oligsub_ref) <- "RNA"
Seurat::Idents(oligsub_ref) <- "MouseID"

# DimPLot
dp <- Seurat::DimPlot(oligsub_ref, reduction = 'umap', group.by = 'TimePoint', alpha = 0.3) |
  Seurat::DimPlot(oligsub_ref, reduction = 'umap', group.by = 'GenoType', alpha = 0.3) |
  Seurat::DimPlot(oligsub_ref, reduction = 'umap', group.by = "MouseID", alpha = 0.3) |
  Seurat::DimPlot(oligsub_ref, reduction = 'umap', group.by = "SampleID", alpha = 0.3)

dp_time <- Seurat::DimPlot(oligsub_ref, reduction = 'umap', group.by = 'TimePoint', alpha = 0.3, split.by = "GenoType")

dp_geno <- Seurat::DimPlot(oligsub_ref, reduction = 'umap', group.by = 'GenoType', alpha = 0.3, split.by = "TimePoint")

dp_tg <- Seurat::DimPlot(oligsub_ref, reduction = 'umap', split.by = 'TimeGeno', alpha = 0.3)

# Save
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_all_refshort.png"), dp, width = 18, height = 4, dpi = 300)
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_timepoint_refshort.png"), dp_time, width = 9, height = 4, dpi = 300)
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_genotype_refshort.png"), dp_geno, width = 9, height = 4, dpi = 300)
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_TimeGeno_refshort.png"), dp_tg, width = 18, height = 4, dpi = 300)
```

## Clustering oligsub
```{r [O] Clustering, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/07_clustering_oligsub.R')

# Read previous object if needed.
# oligsub_man <- qs::qread(file.path(qsave_dir, "a2_olig_processed_manual.qs"))
# oligsub_ref <- qs::qread(file.path(qsave_dir, "a2_olig_processed_refshort.qs"))

oligsub_man <- clustering_oligsub(oligsub_man,
                         range_start = 0.1,
                         range_end = 1.2,
                         range_step = 0.1,
                         save = TRUE,
                         suffix = "_manual",
                         qsave_dir = qsave_dir,
                         markers_dir = file.path(csv_dir, "Markers"))

oligsub_ref <- clustering_oligsub(oligsub_ref,
                         range_start = 0.1,
                         range_end = 1.2,
                         range_step = 0.1,
                         save = TRUE,
                         suffix = "_refshort",
                         qsave_dir = qsave_dir,
                         markers_dir = file.path(csv_dir, "Markers"))

```

### Manual clustering
```{r [V] Check clusters - manual, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object if needed.
# oligsub_man <- qs::qread(file.path(qsave_dir, "a3_clustered_oligsub_manual.qs"))

# Olig markers
olig_markers <- c(
  'Olig1', 'Olig2', 'Sox10', 'Cd9', 'Smarca4', 'Serinc5',
  'Pdgfra', 'Cspg4', 'Lhfpl3', 'Ptprz1', 'Pmp22',
  'Bmp4', 'Enpp6', 'Neu4', 'Prom1',
  'Cnp',
  'Plp1', 'Mbp', 'Mag', 'Enpp4',
  'Mog', 'Cnp', 'Cd82', 'Opalin', 'Klk6', 'Apod', 'Trf',  'Mobp',
  "Tbx18", 'Vtn', 'Lum', 'Col1a2'
  )

# Set assay and Idents
Seurat::DefaultAssay(oligsub_man) <- 'RNA'
Seurat::Idents(oligsub_man) <- oligsub_man$Olig_cluster_0.3

# Glimpse 
Seurat::DimPlot(oligsub_man, reduction = "umap")


# Feature plots
fp_olig <- Seurat::FeaturePlot(oligsub_man, features = olig_markers, label = T)

# Save
ggplot2::ggsave(fp_olig, file = file.path(plot_dir, "FeaturePlot_Olig_markers_manual.png"), width = 16, height= 28, dpi = 300)
```

### ref_short clustering
```{r [V] Check clusters - refshort, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object if needed.
# oligsub_ref <- qs::qread(file.path(qsave_dir, "a3_clustered_oligsub_refshort.qs"))

# Olig markers
olig_markers <- c(
  'Olig1', 'Olig2', 'Sox10', 'Cd9', 'Smarca4', 'Serinc5',
  'Pdgfra', 'Cspg4', 'Lhfpl3', 'Ptprz1', 'Pmp22',
  'Bmp4', 'Enpp6', 'Neu4', 'Prom1',
  'Cnp',
  'Plp1', 'Mbp', 'Mag', 'Enpp4',
  'Mog', 'Cnp', 'Cd82', 'Opalin', 'Klk6', 'Apod', 'Trf',  'Mobp',
  "Tbx18", 'Vtn', 'Lum', 'Col1a2')

# Set assay and Idents
DefaultAssay(oligsub_ref) <- 'RNA'
Idents(oligsub_ref) <- oligsub_ref$Olig_cluster_0.3

# Glimpse 
Seurat::DimPlot(oligsub_ref, reduction = "umap")

# Feature plots
fp_olig <- Seurat::FeaturePlot(oligsub_ref, features = olig_markers, label = T)

# Save
ggplot2::ggsave(fp_olig, file = file.path(plot_dir, "FeaturePlot_Olig_markers_refshort.png"), width = 16, height= 28, dpi = 300)
```

```{r [V] Visualization from paper, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# https://www.sciencedirect.com/science/article/pii/S0959438817301150
# https://www.sciencedirect.com/science/article/pii/S2213671124000778?via%3Dihub
# https://www.sciencedirect.com/science/article/pii/S1534580718305586?via%3Dihub
# https://pmc.ncbi.nlm.nih.gov/articles/PMC5221728/pdf/emss-70326.pdf
# https://onlinelibrary.wiley.com/doi/10.1111/epi.18413
# https://pmc.ncbi.nlm.nih.gov/articles/PMC4691794/

# Set assay and Idents
Seurat::DefaultAssay(oligsub_man) <- "RNA"
Seurat::DefaultAssay(oligsub_ref) <- "RNA"

Seurat::Idents(oligsub_man) <- oligsub_man$Olig_cluster_0.3
Seurat::Idents(oligsub_ref) <- oligsub_ref$Olig_cluster_0.3

# Markers list
markers <- list(
   OligMarkers = c('Olig1', 'Olig2', 'Sox10', 'Cd9', 'Smarca4', 'Serinc5',
                   'Pdgfra', 'Cspg4', 'Lhfpl3', 'Ptprz1', 'Pmp22',
                   'Bmp4', 'Enpp6', 'Neu4', 'Prom1',
                   'Cnp',
                   'Plp1', 'Mbp', 'Mag', 'Enpp4',
                   'Mog', 'Cnp', 'Cd82', 'Opalin', 'Klk6', 'Apod', 'Trf',  'Mobp',
                   "Tbx18", 'Vtn', 'Lum', 'Col1a2'),
  paper1 = c('Ascl1', 'Nfix', 'Dll3', 'Slc38a1', 'Rplp0', 'Rpl4', 'Eef2', 'Npas1', 'Hes', 'Epas1',
             'Pnlip', 'Pcp4', 'Ptprn', 'Nrarp', 'Clu', 'Gcp5', 'Ttyh1'),
  paper2 = c("Top2a", "Fyn", "Etv5", "Ednrb"),
  paper3 = c("Adora1", "Brinp3", "Resp18", "Matn4", "Ccnb1", "Bche", "Cpm", "Elovl7", "Parvb", "Rspo2",
             "Wnt8b", "Clic6", "Trpm3", "Pax2", "Arx", "Lhx6", "Col6a3", "Ddr2", "Abcc9", "Kcnj8",
             "Gjb6", "Ranbp3l", "Itih2", "Il33"),
  paper4 = c("Nr2e1", "Foxg1", "Neurod6", "Otx2", "Nkx2-1", "Eomes", "Hoxa13", "Hoxc11", "Hoxd9",
             "Hoxa11", "Hoxc6", "Hoxb8", "Hoxa3", "Hoxb6", "Hoxc8", "Hoxc9", "Hoxc10", "Hoxa5", "Hoxa6",
             "Hoxd10", "Hoxb2", "Hoxd11", "Hoxa9", "Hoxa7", "Hoxb7", "Hoxa10", "Hoxd9", "Hoxb9", "Pax8",
             "Tlx3", "Nhlh2"),
  paper5 = c("Gli1", "Ebf3", "Scrt2", "Mecom", "Lin28b", "Plag1", "Zic4", "Tcfap2b", "Osr1", "Tead2",
             "Gli3", "Prdm6", "Bnc2", "Gli2", "Smad6", "Hoxb9", "Pax3", "Neurog2", "Lhx9", "Isl1",
             "Neurod6", "Neurod1", "Neurod2", "Dlx1", "Satb2", "Sox8", "Nkx2-2", "Dbx2", "Barx2",
             "Npas1", "Nkx6-2", "Cux1", "Cux2", "Bcl11b"),
  IEGs = c("Jun", "Fos", "Egr1", "Junb")
  )

# FeaturePlot
Seurat::FeaturePlot(oligsub_man, features = markers[['OligMarkers']])
Seurat::FeaturePlot(oligsub_ref, features = markers[['OligMarkers']])

Seurat::FeaturePlot(oligsub_man, features = markers[['paper1']])
Seurat::FeaturePlot(oligsub_ref, features = markers[['paper1']])

Seurat::FeaturePlot(oligsub_man, features = markers[['paper2']])
Seurat::FeaturePlot(oligsub_ref, features = markers[['paper2']])

Seurat::FeaturePlot(oligsub_man, features = markers[['paper3']])
Seurat::FeaturePlot(oligsub_ref, features = markers[['paper3']])

Seurat::FeaturePlot(oligsub_man, features = markers[['paper4']])
Seurat::FeaturePlot(oligsub_ref, features = markers[['paper4']])

Seurat::FeaturePlot(oligsub_man, features = markers[['paper5']])
Seurat::FeaturePlot(oligsub_ref, features = markers[['paper5']])

Seurat::FeaturePlot(oligsub_man, features = markers[['IEGs']])
Seurat::FeaturePlot(oligsub_ref, features = markers[['IEGs']])
```

```{r [O] Name cluster, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Set Idents
Seurat::Idents(oligsub_man) <- oligsub_man$ref_short

# Glimpse
Seurat::DimPlot(oligsub_man, reduction = 'umap', label = T)

# Map
map <- c(
  "0" = "OPC_1", "3" = "OPC_1",
  "5" = "COP",
  "9" = "NFOL",
  "1" = "MFOL",
  "6" = "Mature_OL", "4" = "Mature_OL",
  "2" = "OPC_2", "7" = "OPC_2", "8" = "OPC_2","10" = "OPC_2"
  )

# Add to meta data
oligsub_man@meta.data$OligSub <- map[as.character(oligsub_man$Olig_cluster_0.3)]

# Set Idents
Seurat::Idents(oligsub_man) <- oligsub_man$OligSub

# Glimpse
Seurat::DimPlot(oligsub_man, reduction = 'umap', label = T)

# FeaturePlot
SeuratExtend::FeaturePlot3.grid(oligsub_man, features = c('Pdgfra', 'Cspg4', 'Sox10', 'Enpp6', 'Mog', 'Plp1', 'Mag', 'Mbp', 'Opalin', 'Mobp', 'Serinc5', 'Apod'))

Seurat::FeaturePlot(oligsub_man, features = c('Pdgfra', 'Enpp6', 'Plp1', 'Mag', 'Mbp', 'Apod'), label = T)

# Re-Subset based on manual
oligsub_man2 <- subset(oligsub_man, oligsub_man$OligSub %in% c("OPC_1", "COP", "NFOL", "MFOL", "Mature_OL"))

# Glimpse
Seurat::DimPlot(oligsub_man2, reduction = 'umap', label = T, split.by = "TimeGeno", group.by = "OligSub")
Seurat::FeaturePlot(oligsub_man2, features = c('Pdgfra', 'Enpp6', 'Plp1', 'Mag', 'Mbp', 'Apod'), label = T)

###############################################################
# Set Idents
Seurat::Idents(oligsub_ref) <- oligsub_ref$ref_short

# Glimpse
Seurat::DimPlot(oligsub_ref, reduction = 'umap', label = T)

# FeaturePlot
SeuratExtend::FeaturePlot3.grid(oligsub_ref, features = c('Pdgfra', 'Cspg4', 'Sox10', 'Enpp6', 'Mog', 'Plp1', 'Mag', 'Mbp', 'Opalin', 'Mobp', 'Serinc5', 'Apod'))

Seurat::FeaturePlot(oligsub_ref, features = c('Pdgfra', 'Enpp6', 'Plp1', 'Mag', 'Mbp', 'Apod'), label = T)


```


# (wrong approach- subsettied obj should be processed from integration) Work only with Olig population - Doublet ver.
## Subset Olig only and re-process
```{r [O] Subset, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object
# integrated <- qs::qread(file.path(qsave_dir, "11_dbl_added.qs"))

# Subset based on manual
oligsub_man <- subset(integrated, integrated$CellTypes %in% c("OPC", "NFOL", "MFOL"))

# Subset based on manual
oligsub_ref <- subset(integrated, integrated$ref_short %in% c("OPC", "COP", "NFOL", "MFOL", "Mature_OL"))

# Save
qs::qsave(oligsub_man, file = file.path(qsave_dir, "b1_OligOnlyInit_manual.qs"))
qs::qsave(oligsub_ref, file = file.path(qsave_dir, "b1_OligOnlyInit_refshort.qs"))
```   

```{r [O] Re-process, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object
# oligsub_man <- qs::qread(file.path(qsave_dir, "b1_OligOnlyInit_refshort.qs"))
# oligsub_ref <- qs::qread(file.path(qsave_dir, "b1_OligOnlyInit_manual.qs"))

# Normalize/process 
# check save file name in source code, if you put save = T)
# source("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/06_process_oligsub.R")

# Process integrated object
oligsub_man <- process_oligsub(oligsub_man, save = F)

oligsub_ref <- process_oligsub(oligsub_ref, save = F)

# Save
qs::qsave(oligsub_man, file = file.path(qsave_dir, "b2_olig_processed_manual.qs"))
qs::qsave(oligsub_ref, file = file.path(qsave_dir, "b2_olig_processed_refshort.qs"))
```

## Initial visualization
### Manual cluster
```{r [V] DimPlot manual, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object
# oligsub_man <- qs::qread(file.path(qsave_dir, "b2_olig_processed_manual.qs"))

# set assay and Idents
Seurat::DefaultAssay(oligsub_man) <- "RNA"
Seurat::Idents(oligsub_man) <- "MouseID"

# DimPLot
dp_dbl <- Seurat::DimPlot(oligsub_man, reduction = 'umap', split.by = 'Doublet_Status', alpha = 0.3)

# Save
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_doublet_manual.png"), dp_dbl, width = 18, height = 4, dpi = 300)
```

### ref_short cluster
```{r [V] DimPlot refshort, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object
# oligsub_ref <- qs::qread(file.path(qsave_dir, "b2_olig_processed_refshort.qs"))

# set assay and Idents
Seurat::DefaultAssay(oligsub_ref) <- "RNA"
Seurat::Idents(oligsub_ref) <- "MouseID"

# DimPLot
dp_dbl <- Seurat::DimPlot(oligsub_ref, reduction = 'umap', split.by = 'Doublet_Status', alpha = 0.3)

# Save
ggplot2::ggsave(filename = file.path(plot_dir, "DimPlot_umap_oligsub_doublet_refshort.png"), dp_dbl, width = 18, height = 4, dpi = 300)
```

## Clustering oligsub
```{r [O] Clustering, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# check save file name in source code, if you put save = T)
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/07_clustering_oligsub.R')

# Read previous object if needed.
# oligsub_man <- qs::qread(file.path(qsave_dir, "b2_olig_processed_manual.qs"))
# oligsub_ref <- qs::qread(file.path(qsave_dir, "b2_olig_processed_refshort.qs"))

oligsub_man <- clustering_oligsub(oligsub_man,
                         range_start = 0.1,
                         range_end = 1.2,
                         range_step = 0.1,
                         save = F)

oligsub_ref <- clustering_oligsub(oligsub_ref,
                         range_start = 0.1,
                         range_end = 1.2,
                         range_step = 0.1,
                         save = F)

# Save
qs::qsave(oligsub_man, file = file.path(qsave_dir, "b3_clustered_oligsub_manual.qs"))
qs::qsave(oligsub_ref, file = file.path(qsave_dir, "b3_clustered_oligsub_refshort.qs"))
```


# Proper subset and integration process
## Subsett
```{r [O] Subset Olig, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Get individual RNA objects
processed_list <- qs::qread(file.path(qsave_dir, "04_processed_rna_list_dbl.qs"))

# Subset based on manual
processed_list_man <- processed_list
for (i in seq_along(processed_list_man)) {
  processed_list_man[[i]] <- subset(processed_list_man[[i]], processed_list_man[[i]]$CellTypes %in% c("OPC", "NFOL", "MFOL"))
}

# Subset based on manual
processed_list_ref <- processed_list
for (i in seq_along(processed_list_man)) {
 processed_list_ref[[i]] <- subset(processed_list_man[[i]], processed_list_man[[i]]$ref_short %in% c("OPC", "COP", "NFOL", "MFOL", "Mature_OL"))
}

# Subset based on manual
processed_list_man_npc <- processed_list
for (i in seq_along(processed_list_man_npc)) {
  processed_list_man_npc[[i]] <- subset(processed_list_man_npc[[i]], processed_list_man_npc[[i]]$CellTypes %in% c("OPC", "NFOL", "MFOL", "NPC"))
}

# Subset based on manual
processed_list_ref_npc <- processed_list
for (i in seq_along(processed_list_ref_npc)) {
 processed_list_ref_npc[[i]] <- subset(processed_list_ref_npc[[i]], processed_list_ref_npc[[i]]$ref_short %in% c("OPC", "COP", "NFOL", "MFOL", "Mature_OL", "NPC"))
}

# Save
qs::qsave(processed_list_ref, file = file.path(qsave_dir, "04_processed_rna_list_subset_ref.qs"))
qs::qsave(processed_list_man, file = file.path(qsave_dir, "04_processed_rna_list_subset_man.qs"))
qs::qsave(processed_list_ref_npc, file = file.path(qsave_dir, "04_processed_rna_list_subset_ref_npc.qs"))
qs::qsave(processed_list_man_npc, file = file.path(qsave_dir, "04_processed_rna_list_subset_man_npc.qs"))
```

## Integration
## Integration with RPCA method
Integration with SCT can't cover the number of cells in our object so trying RPCA method in next section
```{r [O] Integrate RNA objects with RPCA, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source code changed for RunPCA due to low cell number in ref obj (P07b has 49 cells). default npcs are 50 and errored out so changed npcs to 48.
# source("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/05__rpca_rna_integration_npc48.R")

# Read previous object if needed.
# processed_list_ref <- qs::qread(file.path(qsave_dir, "04_processed_rna_list_subset_ref.qs"))
# processed_list_man <- qs::qread(file.path(qsave_dir, "04_processed_rna_list_subset_man.qs"))
# processed_list_ref_npc <- qs::qread(file.path(qsave_dir, "04_processed_rna_list_subset_ref_npc.qs"))
# processed_list_man_npc <- qs::qread(file.path(qsave_dir, "04_processed_rna_list_subset_man_npc.qs"))

# Integration
integrated_ref <- rpca_integration(processed_list_ref, 
                               save = F)

integrated_man <- rpca_integration(processed_list_man, 
                               save = F)

integrated_ref_npc <- rpca_integration(processed_list_ref_npc, 
                               save = F)

integrated_man_npc <- rpca_integration(processed_list_man_npc, 
                               save = F)


# Save
qs::qsave(integrated_ref, file = file.path(qsave_dir, "05_integrated_subset_ref.qs"))
qs::qsave(integrated_man, file = file.path(qsave_dir, "05_integrated_subset_man.qs"))
qs::qsave(integrated_ref_npc, file = file.path(qsave_dir, "05_integrated_subset_ref_npc.qs"))
qs::qsave(integrated_man_npc, file = file.path(qsave_dir, "05_integrated_subset_man_npc.qs"))
```

## Process Integrated object
### PCA, Dimentional Reduction
```{r [O] Process integrated object, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/06__process_integrated.R')

# Read previous object if needed.
# integrated_ref <- qs::qread(file.path(qsave_dir, "05_integrated_subset_ref.qs"))
# integrated_man <- qs::qread(file.path(qsave_dir, "05_integrated_subset_man.qs"))
# integrated_ref_npc <- qs::qread(file.path(qsave_dir, "05_integrated_subset_ref_npc.qs"))
# integrated_man_npc <- qs::qread(file.path(qsave_dir, "05_integrated_subset_man_npc.qs"))

# Process integrated object
integrated_ref <- process_integrated(integrated_ref,
                                 save = F)
integrated_man <- process_integrated(integrated_man, 
                               save = F)
integrated_ref_npc <- process_integrated(integrated_ref_npc,
                                 save = F)
integrated_man_npc <- process_integrated(integrated_man_npc, 
                               save = F)
# Save
qs::qsave(integrated_ref, file = file.path(qsave_dir, "06_integrated_subset_ref.qs"))
qs::qsave(integrated_man, file = file.path(qsave_dir, "06_integrated_subset_man.qs"))
qs::qsave(integrated_ref_npc, file = file.path(qsave_dir, "06_integrated_subset_ref_npc.qs"))
qs::qsave(integrated_man_npc, file = file.path(qsave_dir, "06_integrated_subset_man_npc.qs"))
```

### Initial visualization - DimPlot
```{r [O] Add metadata, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Additional meta data
# Forgot to add this early in the process so adding now.
# Convert info list to lookup vectors
timepoint_lookup <- sapply(info, function(x) x[1])
genotype_lookup <- sapply(info, function(x) x[2])

# Add to metadata
integrated_ref@meta.data <- integrated_ref@meta.data %>%
  mutate(
    TimePoint = timepoint_lookup[MouseID],
    GenoType = genotype_lookup[MouseID]
  )

# Add additional meta data combine age and genotype
iintegrated_ref@meta.data$TimeGeno <- paste0(integrated_ref@meta.data$TimePoint, "_", iintegrated_ref@meta.data$GenoType)


# Add to metadata
integrated_man@meta.data <- integrated_man@meta.data %>%
  mutate(
    TimePoint = timepoint_lookup[MouseID],
    GenoType = genotype_lookup[MouseID]
  )

# Add additional meta data combine age and genotype
integrated_man@meta.data$TimeGeno <- paste0(integrated_man@meta.data$TimePoint, "_", integrated_man@meta.data$GenoType)

# Add to metadata
integrated_ref_npc@meta.data <- integrated_ref_npc@meta.data %>%
  mutate(
    TimePoint = timepoint_lookup[MouseID],
    GenoType = genotype_lookup[MouseID]
  )

# Add additional meta data combine age and genotype
integrated_ref_npc@meta.data$TimeGeno <- paste0(integrated_ref_npc@meta.data$TimePoint, "_", integrated_ref_npc@meta.data$GenoType)

# Add to metadata
integrated_man_npc@meta.data <- integrated_man_npc@meta.data %>%
  mutate(
    TimePoint = timepoint_lookup[MouseID],
    GenoType = genotype_lookup[MouseID]
  )

# Add additional meta data combine age and genotype
integrated_man_npc@meta.data$TimeGeno <- paste0(integrated_man_npc@meta.data$TimePoint, "_", integrated_man_npc@meta.data$GenoType)

# Save
qs::qsave(integrated_ref, file = file.path(qsave_dir, "06_integrated_subset_ref.qs"))
qs::qsave(integrated_man, file = file.path(qsave_dir, "06_integrated_subset_man.qs"))
qs::qsave(integrated_ref_npc, file = file.path(qsave_dir, "06_integrated_subset_ref_npc.qs"))
qs::qsave(integrated_man_npc, file = file.path(qsave_dir, "06_integrated_subset_man_npc.qs"))
```

```{r [V] Initial DimPlot after integration, message=TRUE, warning=TRUE, paged.print=TRUE}
# set assay and Idents
Seurat::DefaultAssay(iintegrated_ref) <- "RNA"
Seurat::Idents(iintegrated_ref) <- "MouseID"

Seurat::DefaultAssay(integrated_man) <- "RNA"
Seurat::Idents(integrated_man) <- "MouseID"

Seurat::DefaultAssay(iintegrated_ref_npc) <- "RNA"
Seurat::Idents(iintegrated_ref_npc) <- "ref_short"

Seurat::DefaultAssay(integrated_man_npc) <- "RNA"
Seurat::Idents(integrated_man_npc) <- "CellTypes"

# DimPLot
dp_ref <- Seurat::DimPlot(integrated_ref, reduction = 'umap', group.by = 'TimePoint', alpha = 0.3) |
  Seurat::DimPlot(integrated_ref, reduction = 'umap', group.by = 'GenoType', alpha = 0.3) |
  Seurat::DimPlot(integrated_ref, reduction = 'umap', group.by = "MouseID", alpha = 0.3) |
  Seurat::DimPlot(integrated_ref, reduction = 'umap', group.by = "SampleID", alpha = 0.3)

dp_man <- Seurat::DimPlot(integrated_man, reduction = 'umap', group.by = 'TimePoint', alpha = 0.3) |
  Seurat::DimPlot(integrated_man, reduction = 'umap', group.by = 'GenoType', alpha = 0.3) |
  Seurat::DimPlot(integrated_man, reduction = 'umap', group.by = "MouseID", alpha = 0.3) |
  Seurat::DimPlot(integrated_man, reduction = 'umap', group.by = "SampleID", alpha = 0.3)

dp_ref_npc <- Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', group.by = 'TimePoint', alpha = 0.3) |
  Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', group.by = 'GenoType', alpha = 0.3) |
  Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', group.by = "MouseID", alpha = 0.3) |
  Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', group.by = "SampleID", alpha = 0.3)

dp_man_npc <- Seurat::DimPlot(integrated_man_npc, reduction = 'umap', group.by = 'TimePoint', alpha = 0.3) |
  Seurat::DimPlot(integrated_man_npc, reduction = 'umap', group.by = 'GenoType', alpha = 0.3) |
  Seurat::DimPlot(integrated_man_npc, reduction = 'umap', group.by = "MouseID", alpha = 0.3) |
  Seurat::DimPlot(integrated_man_npc, reduction = 'umap', group.by = "SampleID", alpha = 0.3)

Seurat::DimPlot(integrated_ref, reduction = 'umap', split.by = 'TimePoint', alpha = 0.3)
Seurat::DimPlot(integrated_ref, reduction = 'umap', split.by = 'GenoType', alpha = 0.3)
Seurat::DimPlot(integrated_ref, reduction = 'umap', split.by = "MouseID", alpha = 0.3)
Seurat::DimPlot(integrated_ref, reduction = 'umap', split.by = "SampleID", alpha = 0.3)

Seurat::DimPlot(integrated_man, reduction = 'umap', split.by = 'TimePoint', alpha = 0.3)
Seurat::DimPlot(integrated_man, reduction = 'umap', split.by = 'GenoType', alpha = 0.3)
Seurat::DimPlot(integrated_man, reduction = 'umap', split.by = "MouseID", alpha = 0.3)
Seurat::DimPlot(integrated_man, reduction = 'umap', split.by = "SampleID", alpha = 0.3)

Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', split.by = 'TimePoint', alpha = 0.3)
Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', split.by = 'GenoType', alpha = 0.3)
Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', split.by = "MouseID", alpha = 0.3)
Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', split.by = "SampleID", alpha = 0.3)

Seurat::DimPlot(integrated_man_npc, reduction = 'umap', split.by = 'TimePoint', alpha = 0.3)
Seurat::DimPlot(integrated_man_npc, reduction = 'umap', split.by = 'GenoType', alpha = 0.3)
Seurat::DimPlot(integrated_man_npc, reduction = 'umap', split.by = "MouseID", alpha = 0.3)
Seurat::DimPlot(integrated_man_npc, reduction = 'umap', split.by = "SampleID", alpha = 0.3)
```

## Cluster and Markers, additional meta data
```{r [O] Clusteriazation, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Change save name in source code!!
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/07_clustering_oligsub.R')

# Read previous object if needed.
# integrated_ref <- qs::qread(file.path(qsave_dir, "06_integrated_subset_ref.qs"))
# integrated_man <- qs::qread(file.path(qsave_dir, "06_integrated_subset_man.qs"))
# integrated_ref_npc<- qs::qread(file.path(qsave_dir, "06_integrated_subset_ref_npc.qs"))
# integrated_man_npc <- qs::qread(file.path(qsave_dir, "06_integrated_subset_man_npc.qs"))

# Join layers after integration in RNA assay
Seurat::DefaultAssay(integrated_man) <- "RNA"
integrated_man <- SeuratObject::JoinLayers(integrated_man)

integrated_man <- clustering_oligsub(integrated_man,
                         range_start = 0.4,
                         range_end = 1.2,
                         range_step = 0.1,
                         save = TRUE,
                         suffix = "_man",
                         qsave_dir = qsave_dir,
                         markers_dir = file.path(csv_dir, "Markers"))


# Join layers after integration in RNA assay
Seurat::DefaultAssay(integrated_ref) <- "RNA"
integrated_ref <- SeuratObject::JoinLayers(integrated_ref)

integrated_ref <- clustering_oligsub(integrated_ref,
                         range_start = 0.4,
                         range_end = 1.2,
                         range_step = 0.1,
                         save = TRUE,
                         suffix = "_ref",
                         qsave_dir = qsave_dir,
                         markers_dir = file.path(csv_dir, "Markers"))

# Join layers after integration in RNA assay
Seurat::DefaultAssay(integrated_man_npc) <- "RNA"
integrated_man_npc <- SeuratObject::JoinLayers(integrated_man_npc)

integrated_man_npc <- clustering_oligsub(integrated_man_npc,
                         range_start = 0.4,
                         range_end = 1.2,
                         range_step = 0.1,
                         save = TRUE,
                         suffix = "_man_npc",
                         qsave_dir = qsave_dir,
                         markers_dir = file.path(csv_dir, "Markers"))

# Join layers after integration in RNA assay
Seurat::DefaultAssay(integrated_ref_npc) <- "RNA"
integrated_ref_npc <- SeuratObject::JoinLayers(integrated_ref_npc)

integrated_ref_npc <- clustering_oligsub(integrated_ref_npc,
                         range_start = 0.4,
                         range_end = 1.2,
                         range_step = 0.1,
                         save = TRUE,
                         suffix = "_ref_npc",
                         qsave_dir = qsave_dir,
                         markers_dir = file.path(csv_dir, "Markers"))
```

```{r [V] Glimpse, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
ref <- Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', split.by = 'GenoType', alpha = 0.3, group.by = "ref_short", label = T, repel = T) + NoLegend() + ggtitle("ref") |
Seurat::DimPlot(integrated_ref_npc, reduction = 'umap', split.by = 'GenoType', alpha = 0.3, group.by = "CellTypes", label = T, repel = T) + NoLegend() + ggtitle("manual")

man <- Seurat::DimPlot(integrated_man_npc, reduction = 'umap', split.by = 'GenoType', alpha = 0.3, group.by = "ref_short", label = T, repel = T) + NoLegend() + ggtitle("ref") |
Seurat::DimPlot(integrated_man_npc, reduction = 'umap', split.by = 'GenoType', alpha = 0.3, group.by = "CellTypes", label = T, repel = T) + NoLegend() + ggtitle("manual")

ref / man
```


## Use Mousebrain.org object as reference for cell typing. Using Jesse's custom object.
```{r [O] Cell type, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# source('~/SLC35A2_Olig2cKO_snRNA_JY/WIP/SCRIPTS/SourceCode/08_ref_mousebrain_annotation.R')

# Read previous object if needed.
# integrated_ref <- qs::qread(file.path(qsave_dir, "08_clustered_ref.qs"))
# integrated_man <- qs::qread(file.path(qsave_dir, "08_clustered_man.qs"))
# integrated_ref_npc<- qs::qread(file.path(qsave_dir, "08_clustered_ref_npc.qs"))
# integrated_man_npc <- qs::qread(file.path(qsave_dir, "08_clustered_man_npc.qs"))

# Read reference object for cell typing
ref_obj <- readRDS("~/SLC35A2_Olig2cKO_snRNA_JY/WIP/mouseatlas_cut_processed.rds")

Seurat::DefaultAssay(integrated_ref) <- "SCT"
Seurat::DefaultAssay(integrated_man) <- "SCT"
Seurat::DefaultAssay(integrated_ref_npc) <- "SCT"
Seurat::DefaultAssay(integrated_man_npc) <- "SCT"

# run mouse annotation
integrated_ref <- ref_mousebrain_annotation(ref_obj,
                                        integrated_ref,
                                        save = T,
                                        suffix = "_ref",
                                        save_dir = qsave_dir
                                        )

# run mouse annotation
integrated_man <- ref_mousebrain_annotation(ref_obj,
                                        integrated_man,
                                        save = T,
                                        suffix = "_man",
                                        save_dir = qsave_dir
                                        )

# run mouse annotation
integrated_ref_npc <- ref_mousebrain_annotation(ref_obj,
                                        integrated_ref_npc,
                                        save = T,
                                        suffix = "_ref_npc",
                                        save_dir = qsave_dir
                                        )

# run mouse annotation
integrated_man_npc <- ref_mousebrain_annotation(ref_obj,
                                        integrated_man_npc,
                                        save = T,
                                        suffix = "_man_npc",
                                        save_dir = qsave_dir
                                        )
# Also, cleaned out metadata. all prediction score were deleted.


```

```{r [O] Clean ref annotation, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}
# Read previous object if needed.
# integrated_man <- qs::qread(file.path(qsave_dir, "09_annotated_ref.qs"))

# Ref cell type prediction will be stored in oligsub_man2$predicted.id.
Seurat::Idents(integrated_man) <- integrated_man$predicted.id

# Visualize
Seurat::DimPlot(integrated_man, reduction = "umap", label = T, repel = T) + NoLegend()

# get names and shorten them
ln <- unique(integrated_man$predicted.id)

sn <- c("Ex_Cortex", "Microglia", "Ex_CA3", "NPC", "Ex_CA1", "Interneuron", "Astrocytes", "Granule_DG", "Microglia",
        "MFOL", "Vascular", "Interneuron", "COP", "Interneuron", "Interneuron", "CGE", "Ependymal", "CGE",
        "Mature_OL", "OPC", "Cajal_Retzius", "Interneuron", "Pericytes", "Interneuron", "RGL_DG", "Astrocytes", "Granule_Blast_DG",
        "Macrophages", "Vascular","MGE", "NFOL", "Trilaminar", "Vascular", "Vascular", "Interneuron", "Vascular", "Interneuron")

# Mapping sn-ln
name_map <- setNames(sn, ln)

# Function to replace long names to short names
shortened <- function(name) {ifelse(name %in% names(name_map), name_map[name], name)}

# Apply function
integrated_man$ref_short <- shortened(integrated_man$predicted.id)

# Check UMAP
Seurat::Idents(integrated_man) <- "ref_short"
dp_ref <- Seurat::DimPlot(integrated_man, reduction = "umap", label = T, repel = T) + Seurat::NoLegend()

# Save
ggplot2::ggsave(dp_ref, file = file.path(plot_dir, "DimPlot_mousebrain_ref_sub_ref.png"), width = 8, height = 8, dpi = 300)
qs::qsave(integrated_man, file = file.path(qsave_dir, "09_annotated_ref.qs"))

# Read previous object if needed.
# integrated_man <- qs::qread(file.path(qsave_dir, "09_annotated_man.qs"))

# Ref cell type prediction will be stored in oligsub_man2$predicted.id.
Seurat::Idents(integrated_man) <- integrated_man$predicted.id

# Visualize
Seurat::DimPlot(integrated_man, reduction = "umap", label = T, repel = T) + NoLegend()

# get names and shorten them
ln <- unique(integrated_man$predicted.id)

sn <- c("Ex_Cortex", "Microglia", "Ex_CA3", "NPC", "Ex_CA1", "Interneuron", "Astrocytes", "Granule_DG", "Microglia",
        "MFOL", "Vascular", "Interneuron", "COP", "Interneuron", "Interneuron", "CGE", "Ependymal", "CGE",
        "Mature_OL", "OPC", "Cajal_Retzius", "Interneuron", "Pericytes", "Interneuron", "RGL_DG", "Astrocytes", "Granule_Blast_DG",
        "Macrophages", "Vascular","MGE", "NFOL", "Trilaminar", "Vascular", "Vascular", "Interneuron", "Vascular", "Interneuron")

# Mapping sn-ln
name_map <- setNames(sn, ln)

# Apply function
integrated_man$ref_short <- shortened(integrated_man$predicted.id)

# Check UMAP
Seurat::Idents(integrated_man) <- "ref_short"
dp_ref <- Seurat::DimPlot(integrated_man, reduction = "umap", label = T, repel = T) + Seurat::NoLegend()

# Save
ggplot2::ggsave(dp_ref, file = file.path(plot_dir, "DimPlot_mousebrain_ref_sub_man.png"), width = 8, height = 8, dpi = 300)
qs::qsave(integrated_man, file = file.path(qsave_dir, "09_annotated_man.qs"))
```


```{r [O] Filter low number cell cluster, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

```

```{r [O] DEG, echo=TRUE, message=TRUE, warning=TRUE, paged.print=TRUE}

```

